---
title: "IntroComp_WeixuanChen_HW4"
author: "Weixuan Chen"
date: "2/4/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

## a
The fake grades have a normal distribution with mean 70 and standard deviation 10 from problem 3.d in HW3.
```{r}
my_score <- 45
z_score <- (my_score - 70)/10
z_score
```

## b
The quantile is equivalent to the probability P(X < x). We can see the two methods below are equivalent. The quantile is 0.62%.
```{r}
pnorm(-2.5, mean = 0, sd = 1)

pnorm(45, mean = 70, sd = 10)
```

## c
We have to calculate the probability of two tails. Since normal distribution is symmetric, we can use:
```{r}
2*pnorm(-2.5, mean = 0, sd = 1)
```

## 2

## a

```{r}
set.seed(1)
population <- rnorm(10000, mean = 0, sd = 1)
my_sample <- sample(population, 9, replace = FALSE)
my_sample
```

## b

```{r}
mean(my_sample)
```

${ sum  = -2.1386093 + 0.3926101 - 0.9772168+1.8529964 -0.3512498 -0.3004396+1.9884856+0.2008287 -0.8468195 = -.1794143}$
${sample\:mean =\;\frac{1}{N}\sum_{i = 1}^{N}x_{i}\;=-0.1794143\:/\:9 = -0.01993492}$

## c

```{r}
sd(my_sample)
```
${\ s = \sqrt{\frac{1}{N-1}\sum_{i=1}^{N}(x_i-\bar{x})^2}}$
${s = \sqrt{\frac{1}{9-1}(-2.1386093+0.01993492)^2+(0.3926101-0.01993492)^2+(0.9772168-0.01993492)^2+(1.8529964-0.01993492)^2+(0.3512498-0.01993492)^2 +(0.3004396-0.01993492)^2+(1.9884856-0.01993492)^2+(0.2008287-0.01993492)^2+(0.8468195-0.01993492)^2} = 1.324666}$

## d

```{r}
sd(my_sample)/sqrt(length(my_sample))
```
${se=sd/\sqrt{n}=1.324666/\sqrt{9}=1.324666/3=0.4415553}$

## e

```{r}
mean(my_sample)-1.96*sd(my_sample)/sqrt(length(my_sample))
mean(my_sample)+1.96*sd(my_sample)/sqrt(length(my_sample))
```
${CI = (\bar{x}-1.96*se,\bar{x}+1.96*se) = (-0.8853832, 0.8455134)}$
1.96 here is the z-score for two-sided normal distribution at 97.5% percentile.

## f

```{r}
t <- qt(0.975, 8)
t
mean(my_sample)-t*sd(my_sample)/sqrt(length(my_sample))
mean(my_sample)+t*sd(my_sample)/sqrt(length(my_sample))
```
${CI = (\bar{x}-2.306004*se,\bar{x}+2.306004*se) = (-1.038163, 0.9982933)}$
2.306004 here is the z-score for two-sided t-distribution at 97.5% percentile.

## 3

## a
Since we only have sample size 9, which is less than 30, we cannot use normal distribution to approach the distribution of sample mean. 

## b

```{r}
qnorm(0.95, 0, 1)
qt(0.90, 3)
qt(0.90, 4)
qt(0.95, 3)
qt(0.95, 4)
```
1. It uses the sd of sample, not se of the sample mean, and the degrees of freedom of t-score is 3, not 4, and the quantile should be 95% because it is two-sided, not 90%.
2. The degrees of freedom of t-score is 3, not 4
3. The quantile should be 95% because it is two-sided, not 90%. It is correct if we want one-sided CI, and it uses the sd of sample, not se of the sample mean
4. correct
5. The degrees of freedom of t-score is 3, not 4

## 4

## a
Since the gap between the lower and upper bound of CI is exactly the ${t_{0.975,8}*se = t_{0.975,8}*sd/\sqrt{n}}$. If we want to shrink the CI by 1/2, we at lease have to shrink the gap by 1/2, which means to shrink ${t_{0.975,8}*se = t_{0.975,8}*sd/\sqrt{n}}$ by 1/2. So if we increase the sample size by 4 times, we have ${t_{0.975,8}*se = t_{0.975,8}*sd/\sqrt{4n} = t_{0.975,8}*sd/(2*\sqrt{n}) = 1/2 *t_{0.975,8}*sd/\sqrt{n}}$. So now our sample size is 9, we at least have to increase our sample size to 4*9 = 36, the CI will shrink by 1/2.

## b
Assume that our sample size is greater than 30 (${t-score \approx z-score}$), we have ${Interval = z_{0.975}*sd/(\sqrt{n})}$. ${1.96*20000/\sqrt{n} = 1000\Longrightarrow n =(1.96*20000/1000)^2=1536.64\approx1537}$. 
```{r}
round((1.96*20000/1000)^2)
```

${Interval = z_{0.975}*sd/(\sqrt{n})}$. ${1.96*20000/\sqrt{n} = 100\Longrightarrow n =(1.96*20000/100)^2=153664}$
```{r}
(1.96*20000/100)^2
```
So for CI of ${\pm1000}$, we need at least 1537 people, and for CI of ${\pm100}$, we need at least 153664 people.

## 5
```{r}
# 1. Set how many times we do the whole thing
nruns <- 1000
# 2. Set how many samples to take in each run 
nsamples <- 20
# 3. Create an empty matrix to hold our summary data: the mean and the upper and lower CI bounds.
sample_summary <- matrix(NA,nruns,3)
# 4. Run the loop
counter <- 0
for(j in 1:nruns){
  sampler <- rep(NA,nsamples)
  # 5. Our sampling loop
  for(i in 1:nsamples){
    # 6. At r andom we get either a male or female beetle
    #    If it's male, we draw from the male distribution
    if(runif(1) < 0.5){
      sampler[i] <- runif(n=1,min=5,max=15)
                      }
    #   If it's female, we draw from the female distribution
    else{
      sampler[i] <- runif(n=1,min=15,max=25)
    }
  } 
  # 7. Finally, calculate the mean and 95% CI's for each sample 
  #    and save it in the correct row of our sample_summary matrix
  sample_summary[j,1] <- mean(sampler)  # mean
  standard_error <- sd(sampler)/sqrt(nsamples) # standard error
  sample_summary[j,2] <- mean(sampler) - qt(0.995,19)*standard_error # lower 95% CI bound
  sample_summary[j,3] <- mean(sampler) + qt(0.995,19)*standard_error # lower 95% CI bound
}

counter = 0
for(j in 1:nruns){
  # If 15 is above the lower CI bound and below the upper CI bound:
  if(15 > sample_summary[j,2] && 15 < sample_summary[j,3]){
    counter <- counter + 1
  }
}

print(counter)
```
So our accuracy of CI is 98.5%.
