---
title: "IntroComp_Weixuan_Chen_HW6"
author: "Weixuan Chen"
date: "2/19/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

## a

```{r}
# Create the Tibble
ep <- tibble::tribble(~Party, ~`18-29`, ~`30-44`, ~`45-59`, ~`60+`, "Democrat", 86L, 
    72L, 73L, 71L, "Independent", 52L, 51L, 55L, 54L, "Republican", 61L, 74L, 70L, 
    73L)
# Add sums for cols and rows
ep$Totals <- rowSums(ep[, c(2:5)])
c.sums <- c("Totals", as.vector(colSums(ep[, c(2:6)], dims = 1)))

ep <- rbind(ep, c.sums)
# Apparently rbind converts all the values to characters, this next f(n) converts
# them back to numeric
ep <- transform(ep, `18-29` = as.numeric(`18-29`), `30-44` = as.numeric(`30-44`), 
    `45-59` = as.numeric(`45-59`), `60+` = as.numeric(`60+`))
# Fix the column headers
names(ep) <- c("Party", "18-29", "30-44", "45-59", "60+", "Totals")

ep


```
H0: Age is independent of Party ID.
H1: Party ID and age are dependent variables.
${f_e = (row total)*(column total)/(overall total)}$
${(302)*(199))/(792) = 75.88131}$
${(212)*(199))/(792) = 53.26768}$
${(278)*(199))/(792) = 69.85101}$
${(302)*(197))/(792) = 75.11869}$
${(212)*(197))/(792) = 52.73232}$
${(278)*(197))/(792) = 69.14899}$
${(302)*(199))/(792) = 75.88131}$
${(302)*(198))/(792) = 75.5}$
${(212)*(198))/(792) = 53}$
${(278)*(198))/(792) = 69.5}$
${\chi^2 = \sum{(f_o - f_e)^2/f_e}}$
${\chi^2 = \sum{(86−75.88131)^2/75.88131 = 1.34931539}}$
${\chi^2 = \sum{(52−53.26768)^2/53.26768 = 0.03016847}}$
${\chi^2 = \sum{(61−69.85101)^2/69.85101 = 1.12153539}}$
${\chi^2 = \sum{(72−75.11869)^2/75.11869 = 0.1294779}}$
${\chi^2 = \sum{(51−52.73232)^2/52.73232 = 0.0569090}}$
${\chi^2 = \sum{(74−69.14899)^2/69.14899 = 0.3403130}}$
${\chi^2 = \sum{(73−75.5)^2/75.5 = 0.082781457}}$
${\chi^2 = \sum{(55−53.0)^2/53.0 = 0.075471698}}$
${\chi^2 = \sum{(70−53.0)^2/69.5 = 0.003597122}}$
${\chi^2 = \sum{(71−75.5)^2/75.5 = 0.26821192}}$
${\chi^2 = \sum{(54−53.0)^2/53.0 = 0.01886792}}$
${\chi^2 = \sum{(73−69.5)^2/69.5 = 0.17625899}}$
${\chi^2 = 3.652908}$
${df = (3-1)(4-1) = 6}$
${\chi^2_{critical} = 12.59159}$
${p-value = 0.7235272}$
Conclusion: We cannot reject the null hypothesis.

## 2

```{r}
qchisq(0.95, 6)
```

```{r}
1 - pchisq(3.652908, 6)
```

```{r}
chisq.test(ep[c(1:3), c(2:5)])
```
The answers are consistent

## Problem 2

## a
H0:${\mu_D}$=${\mu_I}$=${\mu_R}$ The average age is the same across the affiliations. H1: The average age is different across the affiliations.

${f_{stat}}$ = (average variance between groups)/(average variance within groups)
between groups = ${n_1(\bar{y_1} - \bar{y}) + ...+ n_G(\bar{y_G} - \bar{y})/df = G-1}$
between groups = ${302(43.3−44.2)^2+212(44.6−44.2)^2+278(45.1−44.2)^2/df = 3-1}$
within groups = ${s_1^2(n_1 - 1) + ...+ s_G^2(n_G - 1)/df = N-G}$
within groups = ${1.28^2(302−1) +1.43^2(212−1)+ 1.1^2(278 - 1)/df = 792-3}$
Let's use R to compute the result.
```{r}
# Summary data input vectors
mu <- c(43.3, 44.6, 45.1)
sd <- c(9.1, 9.2, 9.2)
n <- c(302, 212, 278)

# Function for anova test with input values

# Inputs:

# y - Vector of means

# mu - Mean of means, finds it from input values if not declared

# sd - Vector of standard deviations

# n - Vector of group totals

# Returns - f-statistic

anva <- function(y, mu = mean(y), s, n) {
    wg.v <- vector("numeric")
    bg.v <- vector("numeric")
    for (i in 1:length(n)) {
        # Between Group Variance
        bgc <- n[i] * (y[i] - mu)^2
        bg.v <- append(bg.v, bgc, after = length(bg.v))
        i <- i + 1
    }
    bgvar <- sum(bg.v)/(length(n) - 1)
    i <- 1
    for (i in 1:length(n)) {
        # within Group Variance
        wgc <- (n[i] - 1) * s[i]^2
        wg.v <- append(wg.v, wgc, after = length(wg.v))
        i <- i + 1
    }
    wgvar <- sum(wg.v)/(sum(n) - length(n))
    fs <- bgvar/wgvar
    pv <- 1 - pf(fs, length(n) - 1, (sum(n) - length(n)))
    output <- tibble::tribble(~Param, ~Value, "Fstat", fs, "pValue", pv)
    return(output)
}
anva(mu, 44.2, sd, n)
```
We cannot reject the null hypothesis.

## 2

```{r}
# Create the data frames with rnorm
D <- data.frame(rnorm(n[1], mean = mu[1], sd[1]))
I <- data.frame(rnorm(n[2], mean = mu[2], sd[2]))
R <- data.frame(rnorm(n[3], mean = mu[3], sd[3]))
```

```{r}
D['Indicator'] = "D"
I['Indicator'] = "I"
R['Indicator'] = "R"
colnames(D) <- c("age", "indicator")
colnames(I) <- c("age", "indicator")
colnames(R) <- c("age", "indicator")
```

```{r}
age_df = rbind(D,I,R)
age_df
```

```{r}
aPaov <- aov(age ~ indicator, data = age_df)
summary(aPaov)
```
We can see the result is not consistent with 2a. The possible reason is that the sample we have cannot be a good representative for the population, because we have dataset with different summary statistics. The differences in mean and sd can result in the fluctuation in the values of final statistic. 