---
title: "IntroComp_Weixuan_HW7"
author: "Weixuan Chen"
date: "3/7/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
```

## Data Preparation

```{r}
ageIQ <- tibble::tribble(~Age, ~IQ, 23L, 100L, 18L, 105L, 10L, 95L, 45L, 120L)
ageIQ.s <- cbind(ageIQ, Sums = rowSums(ageIQ))
ageIQ.s <- rbind(ageIQ.s, Sums = colSums(ageIQ.s))
summary(ageIQ)
```

## 1

```{r}
ageIQ
```

```{r}
ggplot(data = ageIQ, mapping = aes(x = Age, y = IQ)) + geom_point() + geom_smooth(method = lm) + 
    ggtitle("Age V IQ") + theme(plot.title = element_text(hjust = 0.5))
```

## 2

${Cov(x,y) = \frac{1}{4-1}\sum{[(23-24)(100-105)+(18-24)(105-105)....+(45-24)(120-105)]}=153.333333}$

```{r}
ageIQ.cov <- ageIQ %>% mutate(S = (Age - mean(ageIQ$Age)) * (IQ - mean(ageIQ$IQ))) %>% 
    colSums()
c(ageIQ.cov[3]/3, cov(ageIQ$Age, ageIQ$IQ))
```

## 3

${r=\frac{153.33333}{14.988*10.80123}=0.947}$

```{r}
(r <- 153.33333/(sd(ageIQ$Age) * sd(ageIQ$IQ)))

c(r, cor(ageIQ$Age, ageIQ$IQ))
```

## 4

${r=\frac{Cov(x,y)}{S_xS_y}=\beta_1\frac{S_x}{S_y}}$
${\beta_1=\frac{rs_y}{s_x}=\frac{0.9470957*10.80123}{14.98888888}=0.6824926}$
${\beta_0=\bar{y}-\beta_1\bar{x}}=105-0.6824926*24=88.62018$
${\hat{y}=88.62018+0.682492x_i}$

```{r}
(b1 <- r/(sd(ageIQ$Age)/sd(ageIQ$IQ)))

(b0 <- mean(ageIQ$IQ) - b1 * mean(ageIQ$Age))
```

## 5

${104.3175=(88.62018)+(0.6824926)23}$
${100.9050=(88.62018)+(0.6824926)18}$
${95.4451=(88.62018)+(0.6824926)10}$
${119.3323=(88.62018)+(0.6824926)45}$

```{r}
(y.i <- ageIQ %>% mutate(`y^i` = b0 + b1 * Age))
```

## 6

${R^2=\frac{TSS-SSE}{TSS}}$
${TSS=\sum{(y_i-\bar{y})^2}=\sum{(100−105)^2+(105−105)^2+(95−105)^2+(120−105)^2=350}}$
${SSE=\sum{(100−104.3175)^2+(105−100.9050)^2+(95−95.4451)^2+(120−119.332)^2}=36.05341}$
${R^2=\frac{TSS-SSE}{TSS}=\frac{350-36.05341}{350}=0.8969903}$

```{r}
(TSS <- ageIQ %>% mutate(tss = ((IQ - mean(ageIQ$IQ))^2)))

sum(TSS$tss)

(SSE <- y.i %>% mutate(sse = (IQ - `y^i`)^2))

sum(SSE$sse)

(`r^2` <- (sum(TSS$tss) - sum(SSE$sse))/sum(TSS$tss))
```
The R^2 value describes the proportion of the variation in the Y data that is explained by the values for X, in other words, how well the line of best fits predicts the data. The R^2 value is also sometimes called the proportional reduction in error and is described as the proportional reduction of error in the variation in Y that would be explained by the line of best fit. A value of 0.897 suggests that ~89.7% of the variation in Y can be explained by X.

## 7

${se_{\hat{y}}=\sqrt{\frac{SSE}{N-2}}=\sqrt{\frac{36.05341}{4-2}}=4.245787}$
${se_{\beta_0}=se_{\hat{y}}\sqrt{\frac{\sum{x_i^2}}{n\sum{(x_i-\bar{x}^2)}}}=4.245787\sqrt{\frac{2978}{4*674}}=0.1635416}$

t-test:
H0:${\beta_1=0}$
H1:${\beta_1\neq0}$

${t_{stat}=\frac{\beta_1-\mu_0}{se_{\beta_1}}=\frac{0.6824926-0}{0.1635416}=4.173205}$
${df=n-k-1=4-1-1=2}$
${t_{crit}=4.302653}$

```{r}
(sey <- sqrt(sum(SSE$sse)/2))

(se <- ageIQ %>% mutate(`xi^2` = Age^2, `xi-x^2` = (Age - mean(ageIQ$Age))^2))

(seb0 <- sey * sqrt(sum(se$`xi^2`)/(4 * sum(se$`xi-x^2`))))

(seb1 <- sey * (1/sqrt(sum(se$`xi-x^2`))))

(ts <- (b1 - 0)/seb1)

qt(0.975, 2)
```
We fail to reject the null hypothesis at the 95% confidence level with a 2-tailed test. The positive correlation between age and IQ is not statistically significant. The failure to reject the null in this t-test of significance indicates that despite our Pearson coeffecient r and proportional reduction in error R^2 suggesting that the linear regression model matches the data with reasonable accuracy, the positive correlation between age and IQ is not significant, and the positive correlation in this instance may be due to chance.

## 8

```{r}
(pv <- 2 * pt(ts, 2, lower.tail = F))
```
The pvalue=0.05290431 suggests that if sample data was gathered on age and IQ with n>100 , about ~5.2% of those trials would have a mean less than or equal to the one in this trial. Since α=.05, we conclude that at the 95% confidence level α/p=.05 with a 2-tailed test, the result for this trial is not statistically significant.

## 9

CI = 0.6824926±4.173205∗0.1635416=[1.110223e−16,1.364985]

```{r}
(ci <- c(b1 + ts * seb1, b1 - ts * seb1))
```
The 95% confidence interval for β1 indicates that for each year of age we would expect that IQ would increase between 1.110223e−16 and 1.364985. One end is very close to 0, suggesting that as each year passes, IQ may not increase hardly at all, and the other end suggesting a gain of up to 1.36 IQ points per year.

## 10

```{r}
summary(lm(IQ ~ Age, data = ageIQ))
```
The results of the lm function agree with the previous calculations.

## 11
The line of best fit and standard error was included in the initial plot.

## 12
r=.947 indicating there is a strong positive correlation between age and IQ. R2=0.897 indicates that approximately 89.7% of the error has been reduced and is explained by the line of best fit. The t-test at the 95% confidence level yielding a p-value of  .053 indicates that despite the strength of the correlation indicated by the r and R2 values, this correlation is not signficant, may not actually be as positive as the line of best fit in this trial suggests.

