---
title: "IntroComp_Weixuan_HW5"
author: "Weixuan Chen"
date: "2/12/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

4.
Using data of your choosing (or using simulated data), use R to conduct the following tests, and explain the results you get:

a. A standard one-sample hypothesis test.
```{r}
datasets::cars
t.test(cars$dist,alternative="two.sided",mu=60)
```
Here, in the cars dataset, we have, Null Hypothesis $H_{0}: \mu=60$
Since, p value < alpha (0.05), we reject the null hypothesis.



b. A difference-in-means test with independent samples.

```{r}
s1 <- sample(15:45, 10)
s2 <- sample(20:50, 10)
t.test(s1,s2,mu=0,conf=0.95,alternative="two.sided")
```
Here, the Null hypothesis, $H_{0}:\mu=0$
However, since the p-value = 0.0539, we cannot reject the null hypothesis at 0.05 level.



c. A difference-in-means test with dependent samples (ie., a paired t-test)

Here, we'll use an example data set, which contains the weight of 10 mice before and after the treatment.
```{r}
# Weight of the mice before treatment
before <-c(200.1, 190.9, 192.7, 213, 241.4, 196.9, 172.2, 185.5, 205.2, 193.7)
# Weight of the mice after treatment
after <-c(392.9, 393.2, 345.1, 393, 434, 427.9, 422, 383.9, 392.3, 352.2)
# Compute t-test
t.test(before, after, paired = TRUE)
```
As we can see that the p value is significantly less than alpha,
hence we reject Null hypothesis.



d. Manually verify the results in (a) using the mean and sd as calculated by R (ie, you don't have to
manually calculate the mean or sd by hand!).

```{r}
x_bar <- mean(cars$dist)
x_bar
```

```{r}
std_dev <- sd(cars$dist)
std_dev
```

```{r}
n <- length(cars$dist)
n
```
```{r}
se <- std_dev/(sqrt(n))
se
```
```{r}
mu <- 60
z <- (x_bar - mu)/se
z
```
```{r}
low_int <- qt(0.025,n-1)
low_int
high_int <-qt(0.975,n-1)
high_int
```
```{r}
low_CI <- ((low_int)*se)+x_bar
low_CI
high_CI <- ((high_int)*se)+x_bar
high_CI
```
This verifies the results from part a and Null hypothesis can be rejected since mu lies in the rejected area.


